AFQ-Insight

### 2019-12-03 reports/manuscript/abstract.tex
Add details to cross-validation section and software implementation. Update some references and minor edits to other sections

The white matter contains long-range connections between different brain
regions. Tractometry uses diffusion-weighted magnetic resonance imaging (dMRI)
data to quantify tissue properties, such as fractional anisotropy (FA) and mean
diffusivity (MD), along the trajectories of these connections in the human brain
\emph{in vivo}\cite{yeatman2012tract}. Statistical inference from tractometry
data usually either averages these quantities along the length of the tract in
each individual, or performs analysis point-by-point along each tract, with
group comparisons or regression models computed separately for each point along
every one of the tracts. Alternatively, tissue properties are computed for a
specific tract of interest based on an a priori hypothesis. In the present work,
we developed a method based on the sparse group lasso (SGL)
\cite{simon2013sparse} that takes into account tissue properties measured along
the tracts, and selects informative features by enforcing sparsity, both at the
level of individual tracts and tissue properties, but also across the entire set
of tracts and all of the measured tissue properties. The sparsity penalties for
each of these constraints is identified using a nested cross-validation scheme
that guards against over-fitting and simultaneously identifies the correct level
of sparsity, both at the group level and overall. We demonstrate the accuracy of
the method in two settings: In a regression setting, dMRI is used to accurately
predict ``brain age'' \cite{yeatman2014lifespan, Brown2012-so}. In a
classification setting, patients with amyotrophic lateral sclerosis (ALS) are
accurately distinguished from matched controls\cite{sarica2017corticospinal}. In
addition, SGL automatically identifies FA in the corticospinal tract as
important for this classification -- correctly finding the parts of the white
matter known to be affected by ALS. \ariel{Still need a punchy ending sentence
here}

>>>>

The white matter contains long-range connections between different brain
regions. Tractometry uses diffusion-weighted magnetic resonance imaging (dMRI)
data to quantify tissue properties, such as fractional anisotropy (FA) and mean
diffusivity (MD), along the trajectories of these connections in the human brain
\emph{in vivo}\cite{yeatman2012tract}. Statistical inference from tractometry
data usually either averages these quantities along the length of the tract in
each individual, or performs analysis point-by-point along each tract, with
group comparisons or regression models computed separately for each point along
every one of the tracts. Alternatively, tissue properties are computed for a
specific tract of interest based on an a priori hypothesis. In the present work,
we developed a method based on the sparse group lasso (SGL)
\cite{simon2013sparse} that takes into account tissue properties measured along
the tracts, and selects informative features by enforcing sparsity, both at the
level of individual tracts and tissue properties, but also across the entire set
of tracts and all of the measured tissue properties. The sparsity penalties for
each of these constraints is identified using a nested cross-validation scheme
that guards against over-fitting and simultaneously identifies the correct level
of sparsity, both at the group level and overall. We demonstrate the accuracy of
the method in two settings: in a regression setting, dMRI is used to accurately
predict ``brain age'' \cite{yeatman2014lifespan, Brown2012-so}. In a
classification setting, patients with amyotrophic lateral sclerosis (ALS) are
accurately distinguished from matched controls\cite{sarica2017corticospinal}. In
addition, SGL automatically identifies FA in the corticospinal tract as
important for this classification -- correctly finding the parts of the white
matter known to be affected by ALS. \ariel{Still need a punchy ending sentence
here}

### 2019-12-03 reports/manuscript/sections/methods.tex
More references and edits throughout.
Add details to cross-validation section and software implementation. Update some references and minor edits to other sections

TODO:

\begin{itemize}
  \item Computational implementation
    \begin{itemize}
      \item Insert Figure for pipeline
      \item Proximal gradient methods
      \item Meta-parameter optimization
      \item Cross-validation scheme
        \begin{itemize}
          \item Insert Figure for cross-validation scheme
        \end{itemize}
    \end{itemize}
\end{itemize}

@@

\item Diffusion MRI data from a previous study of properties of the white matter
across the lifespan\cite{yeatman2014lifespan}, containing dMRI data from 76
subjects with ages 6-50. These data were measured in a GE Discovery 750 3T MRI
scanner at the Stanford Center for Cognitive and Neurobiological Imaging. The
Stanford University IRB approved the procedures of this study. Informed consent
was obtained from each adult participant, and assent for participation was
provided by by parents/guardians for children. Voxel resolution was 2x2x2 $mm^3$
96 non-colinear directions of measurement were measured with a b=2,000 $s/mm^2$
and 30 non-colinear directions of measurement were measured with a b=1,000
$s/mm^2$. Data was preprocessed to correct for subject motion and the diffusion
tensor model \cite{basser1994mr} was fit in every voxel, using a robust fit
\cite{chang2005restore}. These data were acquired using a dual spin echo
sequence, in which there is sufficient time for eddy currents to subside between
the application of the gradients and the image acquisition, so no eddy current
correction was appliced. We will refer to this dataset as WH.

\item Diffusion MRI from from a previous study of the corticospinal tract
profile in patients with amyotrophic lateral sclerosis
(ALS)\cite{sarica2017corticospinal}, containing data from 24 ALS patients and 24
demographically matched healthy controls. These data were measured in a GE
Discovery 750 3T MRI scanner at the Institute of Bioimaging and Molecular
Physiology in Catanzaro. Informed consent was provided as approved by the
Ethical Committee of the University ``Magna Graecia'' of Catanzaro. Voxel
resolution was 2x2x2 $mm^3$ and 27 non-colinear directions were measured with a
b=1,000 $s/mm^2$. Here, both motion correction and eddy current correction were
applied before modeling the diffusioon tensor in each voxel. We will refer to
this dataset as ALS.

@@

where $y$ is the phenotype -- categorical, such as a clinical diagnosis, or
continuous numerical, such as the subject's age. The tractometry data is
represented in the feature matrix $\mathbf{X}$, with rows corresponding to
different subjects, and columns corresponding to diffusion measures at different
nodes within each tract. The relationship between tractometric features and the
phenotypic target is characterized by the coefficients in $\beta$. In general,
the feature matrix $\mathbf{X}$ has dimensions $S \times (T \cdot N \cdot M)$,
where $S$ is the number of subjects, $T$ is the number of white matter tracts,
$N$ is the number of nodes in each tract, and $M$ is the number of diffusion
metrics calculated at each node. Typically, $T = 20$, $N = 100$, and $2 \le M
\le 8$, resulting in $\sim 4,000 - 16,000$  features. Conversely, many dMRI
studies have between a few dozen and a few hundred subjects, yielding a feature
matrix that is wide and short. Even in cases where more than a thousand subjects
are measured (e.g., in the Human Connectome Project, where 1,200 subjects were
measured \cite{VanEssen2012}), the problem is ill-posed: The high dimensionality
of this data requires regularization to avoid overfitting and generate
interpretable results.

Here, we propose that in addition to regularizing the coefficients in $\beta$,
we can also capitalize on our knowledge of the group structure of the tract
profile features in $\mathbf{X}$. The tract-metric combinations form a natural
grouping. For example, we expect that MD features within the left arcuate
fasciculus to co-vary across individuals. Likewise for FA values within the
right corticospinal tract (CST) and so on. This group structure is represented
in Fig~\ref{fig:group-structure}, which depicts the linear model $y = \mathbf{X}
\cdot \beta$. Thus, we seek a regularization approach that will fit a linear
model with anatomically-grouped covariates, where we expect to observe both
groupwise sparsity: the number of groups (tract/metric combinations) with at
least one non-zero coefficients as well as within-group sparsity: the number of
non-zero coefficients within each non-zero groups. The sparse group lasso (SGL)
is a penalized regression technique that satisfies exactly these
criteria\cite{simon2013sparse}. It solves for a coefficient vector
$\widehat{\beta}$ that satisfies
\begin{equation}
    \widehat{\beta} = \min_\beta \frac{1}{2}
    \norm*{y - \displaystyle \sum_{\ell = 1}^{G} \mathbf{X}^{(\ell)} \cdot \beta^{(\ell)}}_2^2
    + \lambda_1 \displaystyle \sum_{\ell = 1}^{G} \sqrt{p_\ell} \norm*{\beta^{(\ell)}}_2
    + \lambda_2 \norm*{\beta}_1,
    \label{eq:sgl}
\end{equation}

where $G$ is the number of groups $\mathbf{X}^{(\ell)}$ is the submatrix of
$\mathbf{X}$ corresponding to group $\ell$, $\beta^{(\ell)}$ is the coefficient
vector for group $\ell$ and $p_\ell$ is the length of $\beta^{(\ell)}$. In the
tractomtetry setting, $G = T \cdot M$ and $\forall \ell: p_\ell = 100$. The
first term is the mean square error loss, $L_{mse}$, as in the standard linear
regression framework. The second and third terms encourage groupwise sparsity
and overall sparsity, respectively. If $\lambda_1 = 0$ and $\lambda_2 = 1$, the
SGL reduces to the traditional lasso\cite{tibshirani1996regression}. Conversely,
if $\lambda_1 = 1$ and $\lambda_2 = 0$, the SGL reduces to the group
lasso\cite{yuan2006model}.

@@

For given values of $\lambda_1$ and $\lambda_2$, the cost function in equation
\ref{eq:sgl} can be optimized using proximal gradient descent methods
\cite{parikh2014proximal} here implemented as a custom proximal operator that is
then optimized using the COPT library \ariel{citation needed here}. To
find the values of the metaparameters $\lambda_1$ and $\lambda_2$, we used a
hyperoptimization procedure, implemented using the hyperopt software
\cite{Bergstra_2015}. XXX We need more details here.

To objectively evaluate the model and guard against over-fitting, we used
cross-validation. XXX More here.

@@

Often, the target variable $y$ is not in the domain in which the linear model
can be best fit to it. Equation \ref{eq:lm} can be slightly modified as:

\begin{equation}
y = f^-1(\mathbf{X} \cdot \beta),
\label{eq:lm}
\end{equation}

where the transformation function $f^-1$ characterizes the transform applied to
the data before fitting the linear coefficients. For example, an often-used
transform is the logarithmic transform:

\begin{equation}
f(y) = log_n(y)
\label{eq:log_nonlinearity}
\end{equation}

In this case, the model is parametrized by one additional fit parameter, $n$.

\subsubsection*{Classification of categorical $y$}
When the phenotypical target variable is categorical, as in the case of
explaining or predicting the presence of a clinical diagnosis, the SGL is
readily adapted to logistic regression, where the probability of a target
variable belonging to an arbitrary defined ``true'' class is the logistic
function of the result of the linear model,

\begin{equation}
    p(y = 1) = \frac{1}{1 + \exp(\mathbf{X}\cdot \beta)},
    \label{eq:logit}
\end{equation}

or equivalently, the mean squared error loss function in Eq~\eqref{eq:sgl} is
replaced with the cross-entropy loss, which for binary classification is the
negative log likelihood of the SGL classifier giving the ``true'' label:

\begin{equation}
    L_{mse} \rightarrow L_{\log} = -\left(y \log(p) + (1 - y) \log(1 - p)\right).
    \label{eq:logloss}
\end{equation}

@@

The full software implementation of the SGL approach presented here is available
in a Python software package called AFQ-Insight, which is developed publicly in
\url{https://github.com/richford/afq-insight}. The version of the code used to
produce the results herein is also available in XXX Zenodo DOI. AFQ-Insight
reads the target and feature data that has been processed by AFQ from comma
separated value (CSV) files conforming to the AFQ-Browser data
format\cite{yeatman2018browser} and represents them internally as
\lstinline{DataFrame} objects from the pandas Python
library\cite{mckinney2010data}. The software provides different options for
imputing missing data in the feature matrix. Missing interior nodes are imputed
using linear interpolation. For missing exterior nodes, the user may choose
between linear extrapolation and constant forward(back)-fill. Imputation uses
only values from adjacent nodes in the same white matter tract in the same
subject so there is no danger of data leakage from other subjects.

>>>>

@@

\item Diffusion MRI data from a previous study of properties of the white matter
across the lifespan\cite{yeatman2014lifespan}, containing dMRI data from 76
subjects with ages 6-50. These data were measured in a GE Discovery 750 3T MRI
scanner at the Stanford Center for Cognitive and Neurobiological Imaging. The
Stanford University IRB approved the procedures of this study. Informed consent
was obtained from each adult participant, and assent for participation was
provided by parents/guardians for children. Voxel resolution was
\num{2x2x2}\si{\milli\meter^3} with 96 non-colinear directions measured with a
$b= \SI{2000}{\second\per\milli\meter^2}$ and 30 non-colinear directions
measured with a $b=\SI{1000}{\second\per\milli\meter^2}$.
Data was preprocessed to correct for subject motion and the diffusion
tensor model \cite{basser1994mr} was fit in every voxel, using a robust fit
\cite{chang2005restore}. These data were acquired using a dual spin echo
sequence, in which there is sufficient time for eddy currents to subside between
the application of the gradients and the image acquisition, so no eddy current
correction was appliced. We will refer to this dataset as WH.

\item Diffusion MRI from from a previous study of the corticospinal tract
profile in patients with amyotrophic lateral sclerosis
(ALS)\cite{sarica2017corticospinal}, containing data from 24 ALS patients and 24
demographically matched healthy controls. These data were measured in a GE
Discovery 750 3T MRI scanner at the Institute of Bioimaging and Molecular
Physiology in Catanzaro. Informed consent was provided as approved by the
Ethical Committee of the University ``Magna Graecia'' of Catanzaro. Voxel
resolution was \num{2x2x2}\si{\milli\meter^3} and 27 non-colinear directions
were measured with a $b=\SI{1000}{\second\per\milli\meter^2}$.
Here, both motion correction and eddy current correction were
applied before modeling the diffusioon tensor in each voxel. We will refer to
this dataset as ALS.

@@

where $y$ is the phenotype -- categorical, such as a clinical diagnosis, or
continuous numerical, such as the subject's age. The tractometry data is
represented in the feature matrix $\mathbf{X}$, with rows corresponding to
different subjects, and columns corresponding to diffusion measures at different
nodes within each tract. The relationship between tractometric features and the
phenotypic target is characterized by the coefficients in $\beta$. In general,
the feature matrix $\mathbf{X}$ has dimensions $S \times (T \cdot N \cdot M)$,
where $S$ is the number of subjects, $T$ is the number of white matter tracts,
$N$ is the number of nodes in each tract, and $M$ is the number of diffusion
metrics calculated at each node. Typically, $T = 20$, $N = 100$, and $2 \le M
\le 8$, resulting in $\sim 4,000 - 16,000$  features. Conversely, many dMRI
studies have between a few dozen and a few hundred subjects, yielding a feature
matrix that is wide and short. Even in cases where more than a thousand subjects
are measured (e.g., in the Human Connectome Project, where 1,200 subjects were
measured \cite{VanEssen2012}), the problem is ill-posed: the high dimensionality
of this data requires regularization to avoid overfitting and generate
interpretable results.

Here, we propose that in addition to regularizing the coefficients in $\beta$,
we can also capitalize on our knowledge of the group structure of the tract
profile features in $\mathbf{X}$. The tract-metric combinations form a natural
grouping. For example, we expect that MD features within the left arcuate
fasciculus will co-vary across individuals. Likewise for FA values within the
right corticospinal tract (CST) and so on. This group structure is represented
in Fig~\ref{fig:group-structure}, which depicts the linear model $y = \mathbf{X}
\cdot \beta$. Thus, we seek a regularization approach that will fit a linear
model with anatomically-grouped covariates, where we expect to observe both
groupwise sparsity, where the number of groups (tract/metric combinations) with
at least one non-zero coefficients is small, as well as within-group sparsity,
where the number of non-zero coefficients within each non-zero group is small.
The sparse group lasso (SGL)
is a penalized regression technique that satisfies exactly these
criteria\cite{simon2013sparse}. It solves for a coefficient vector
$\widehat{\beta}$ that satisfies
\begin{equation}
    \widehat{\beta} = \min_\beta \frac{1}{2}
    \norm*{y - \displaystyle \sum_{\ell = 1}^{G}
    \mathbf{X}^{(\ell)} \cdot \beta^{(\ell)}}_2^2
    + \lambda_1 \displaystyle \sum_{\ell = 1}^{G}
    \sqrt{p_\ell} \norm*{\beta^{(\ell)}}_2
    + \lambda_2 \norm*{\beta}_1,
    \label{eq:sgl}
\end{equation}

where $G$ is the number of groups $\mathbf{X}^{(\ell)}$ is the submatrix
of $\mathbf{X}$ corresponding to group $\ell$, $\beta^{(\ell)}$ is
the coefficient vector for group $\ell$ and $p_\ell$ is the length of
$\beta^{(\ell)}$. In the tractomtetry setting, $G = T \cdot M$ and
$\forall \ell: p_\ell = 100$. The first term is the mean square error
loss, $L_{\text{mse}}$, as in the standard linear regression framework.
The second and third terms encourage groupwise sparsity and overall
sparsity, respectively. If $\lambda_1 = 0$ and $\lambda_2 = 1$, the
SGL reduces to the traditional lasso\cite{tibshirani1996regression}.
Conversely, if $\lambda_1 = 1$ and $\lambda_2 = 0$, the SGL reduces to
the group lasso\cite{yuan2006model}.

@@

For given values of $\lambda_1$ and $\lambda_2$, the cost function in equation
\ref{eq:sgl} can be optimized using proximal gradient descent methods
\cite{parikh2014proximal} here implemented as a custom proximal operator that is
then optimized using the C-OPT library\cite{copt}. This supplies an estimate
of the optimal $\widehat{\beta}$ given a particular set of values for the
regulation parameters $\lambda_1$ and $\lambda_2$. We then find the optimal
values for $\lambda_1$ and $\lambda_2$ using the hyperopt library's \verb|fmin|
function\cite{Bergstra_2015} with a tree of Parzen estimators search
algorithm\cite{bergstra2011algorithms}. For the WH dataset, where the target
variable is continuous numerical, \verb|fmin| searches for metaparameters that
minimize the median absolute error, though in general a user may also
choose to minimize the root-mean-square-error or to maximize the
coefficient of determination ($R^2$). For the ALS dataset, where the
target variable is binary categorical, \verb|fmin| seeks to maximize the
classification accuracy, though in general the user may also choose to
maximize the area under the receiver operating curve (ROC AUC) or the
average precision. \adam{Enough detail provided?}

To objectively evaluate the model and guard against over-fitting,
we used a nested cross-validation scheme, depicted in
Fig~\ref{fig:cross-val} for the categorical classification case. The
input data is shuffled and then decomposed into $k$ batches, hereafter
referred to as folds. For the ALS dataset, $k=10$, while for the WH
dataset, $k=5$. For each unique fold, we hold that fold out as the
test\textsubscript{outer} set and let the remaining data comprise
the train\textsubscript{outer} set, with the subscript indicating
the depth of the nested decomposition. We further decompose each
train\textsubscript{outer} set into three folds, and again for each
unique fold, we hold out that fold as the test\textsubscript{inner}
set and let the remaining data comprise the train\textsubscript{inner}
set. At level-1 of the decomposition, we fit an SGL model using
arbitrary fixed regularization parameters $\lambda_1$ and $\lambda_2$,
training the model using train\textsubscript{inner} and evaluating
the fit using the appropriate metric on test\textsubscript{inner}.
The optimal regularization parameters are determined using the
aforementioned hyperopt library at level-0 of the decomposition. We
find optimal regularization parameters and $\widehat{\beta}$ for each
train\textsubscript{outer} set and then use those to predict values for
data in test\textsubscript{outer}. Thus each subject in the dataset has
a predicted phenotype derived from a model that never saw its particular
subject's data.

The above procedure describes $k$-fold cross validation. In fact,
we use repeated $k$-fold cross validation on the outer level of the
decomposition, so that the input data is decomposed into $k$ folds,
three times. Thus, each subject has three predicted phenotypes. We then
take the mean predicted value to summarize the skill of the model.
In the classification case, the shuffling into folds is stratified such
that each fold has a population that preserves the percentage of each
class found in the larger input data.
\adam{Clear as mud?}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.55\textwidth]{nested-cross-validation.pdf}
    \caption{{\bf Nested $k$-fold cross-validation scheme.}
        We evaluate model quality using a nested $k$-fold cross
        validation scheme. At level-0, the input data is decomposed
        three times into $k$ shuffled groups and optimal hyperparameters
        are found for the level-0 training set. Optimization of these
        hyperparameters requires the use of the hyperopt library and
        many repeated evaluations of an SGL model over a search space
        of possible regularization parameters. These evaluations take
        place at level-1 of the decomposition, where the level-0 training
        set is further decomposed into three shuffled groups.
        For the ALS data, $k=10$. For the WH data, $k=5$.
    }
    \label{fig:cross-val}
\end{figure}

@@

Often, the target variable $y$ is not in the domain in which the linear model
can be best fit to it. Equation \ref{eq:lm} can be slightly modified as:
\begin{equation}
y = f^-1(\mathbf{X} \cdot \beta),
\label{eq:lm}
\end{equation}
where the transformation function $f^{-1}$ characterizes the transform applied
to the data before fitting the linear coefficients. For example, an often-used
transform is the logarithmic transform:
\begin{equation}
f(y) = \log_n(y)
\label{eq:log_nonlinearity}
\end{equation}
In this case, the model is parametrized by one additional fit parameter, $n$.

\subsubsection*{Classification of categorical $y$}
When the phenotypical target variable is categorical, as in the case of
explaining or predicting the presence of a clinical diagnosis, the SGL is
readily adapted to logistic regression, where the probability of a target
variable belonging to an arbitrary defined ``true'' class is the logistic
function of the result of the linear model,
\begin{equation}
    p(y = 1) = \frac{1}{1 + \exp(\mathbf{X}\cdot \beta)},
    \label{eq:logit}
\end{equation}
or equivalently, the mean squared error loss function in Eq~\eqref{eq:sgl} is
replaced with the cross-entropy loss, which for binary classification is the
negative log likelihood of the SGL classifier giving the ``true'' label:
\begin{equation}
    L_{\text{mse}} \rightarrow L_{\log} =
    -\left(y \log(p) + (1 - y) \log(1 - p)\right).
    \label{eq:logloss}
\end{equation}

@@

The full software implementation of the SGL approach presented here is available
in a Python software package called AFQ-Insight, which is developed publicly in
\url{https://github.com/richford/afq-insight}. The version of the code used to
produce the results herein is also available in \ariel{Need to add Zenodo DOI}.
AFQ-Insight reads the target and feature data that has been processed by AFQ
from comma separated value (CSV) files conforming to the AFQ-Browser data
format\cite{yeatman2018browser} and represents them internally as
\lstinline{DataFrame} objects from the pandas Python
library\cite{mckinney2010data}. The software provides different options for
imputing missing data in the feature matrix. Missing interior nodes are imputed
using linear interpolation. For missing exterior nodes, the user may choose
between linear extrapolation and constant forward(back)-fill. Imputation uses
only values from adjacent nodes in the same white matter tract in the same
subject so there is no danger of data leakage from other subjects. It uses the
scikit-learn\cite{scikit-learn} library to decompose input data into separate
test and train datasets, to scale each feature to have zero mean and
unit variance, and to evaluate each fit in the hyperparameter search using
appropriate classification and regression metrics such as accuracy, area
under the receiver operating curve (AUC ROC), and coefficient of determination
($R^2$). For each set of hyperparameters, we solve the SGL using a custom
proximal operator supplied to the C-OPT library\cite{copt}. Appropriate
hyperparameters are found using the hyperopt library\cite{Bergstra_2015}.

### 2019-12-03 reports/manuscript/sections/results.tex
Add details to cross-validation section and software implementation. Update some references and minor edits to other sections

To test the performance of SGL in tractometry data in a continuous regression
task, we focus here on the prediction of biological age from the tractometry
data. Prediction of ``brain age'' is a commonly undertaken task. This is both
because it operates on a natural scale, with meaningful and easily understood
units, as well as because predictions of brain age, and deviations from accurate
prediction are diagnostic of overall brain health (for a recent review, see
\cite{Cole2019-rz}). The WH dataset used here contains data from 76 healthy
subjects, ranging between 6 years and 50 years of age
\cite{yeatman2014lifespan}. In this case, biological age was used as the
predicted variable ($y$ in equation \ref{eq:lm}). SGL was fit to
tractometry-extracted features: FA and MD in 20 major brain tracts, with each
tract divided into 100 nodes. To evaluate the fit of the model, we used a nested
cross-validation procedure. In this procedure, batches of subjects are held out.
For each batch (or fold), the model is fully fit without this data. Then, once
the parameters are fixed, the model is inverted to predict the ages of held out
subjects based on the linear coeffiecients and the static non-linearity. data is
held out, and a model age was predicted using a model that was fit without that
subject. This scheme automatically finds the right level of regularization and
fits the coefficients to the ill-posed linear model, while guarding against
overfitting. SGL accurately predicts the age of the subjects in this procedure,
with a mean absolute error of 3.6 years (Figure \ref{fig:regress-results}, left
panel). This is lower than the results of a recent study that predicted age in a
large sample, based on diffusion MRI features \cite{Richard2018-ux}.
Nevertheless, older subjects have higher residual variance, reflecting the
automatically-chosen log-transformation and implying that brain age becomes more
difficult to predict as we age chronologically (\ref{fig:regress-results}, right
panel). The model weights are distributed over many different tracts and dMRI
tissue properties (Figure \ref{fig:regress-beta} left). This demonstrates that
SGL is not coerced to produce overly sparse results when a more accurate model
requires a dense selection of features. Furthermore, looking closer at a
selection of tracts where high coefficients are found demonstrates that
diffusion properties (FA, in this case) are different in different age groups in
parts of the tracts where these higher coefficients are found (Figure
\ref{fig:regress-beta} right).

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{regression_residuals.png}
    \caption{{\bf Predicting age with tractometry and SGL.} Left: The predicted
    age of each individual (on the abscissa) and true age (on the ordinate),
    from the test splits (i.e., when each subject's data was held out in fitting
    the model); an accurate prediction falls close to the $y=x$ line (dashed).
    The mean absolute error in this case is 3.6 years and, the coefficient of
    determination $R^2=0.3$. Right: Standardized residuals (on the abscissa) as
    a function of the true age (on the ordinate). Predictions are generally more
    accurate for younger individuals.}
    \label{fig:regress-results}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.3\textwidth]{regression_beta_annotated.pdf}
    \includegraphics[width=0.67\textwidth]{regression_tract_profiles.pdf}
    \caption{{\bf Feature importance for predicting age from tractometry.} Left:
    A skeletonized display of the main brain tracts analyzed, with anterior
    facing up, and right hemisphere on the right. The $\hat{\beta}$ coefficients
    displayed in blue (negative) to red (positive) are for measurements of FA
    along the length of the tracts. The left cingulum cingulate (A) and forceps
    minor (B) are highlighted. Right: the FA (in shades of blue and green) and
    the $beta$ coefficients (dashed) in (A) left cingulum and (B) forceps
    minor.}
    \label{fig:regress-beta}
\end{figure}

@@

Using data from a previous study of the corticospinal tract (CST) profile and
ALS\cite{sarica2017corticospinal}, we tested the performance of SGL in a
classification setting. The previous study predicted ALS status with a mean
accuracy of 80\% using a random forest algorithm based on a priori selection of
features within the corticospinal tract. SGL delivers competitive predictive
performance (mean 93\% $\pm$ 2\% accuracy, 0.978 $\pm$ 0.006 ROC AUC) without
the need for a priori feature engineering. The results of the classification
prediction are shown in Fig~\ref{fig:class-results}, with ``ground-truth'' ALS
status separated into columns, and predicted ALS status encoded by color. In
addition to this classification performance, SGL also identifies the white
matter tracts most important for ALS classification. The relative importance of
white matter features is captured in the $\beta$ coefficients from
Eq~\eqref{eq:sgl}. Fig~\ref{fig:class-beta} depicts these coefficients across
the brain, laid out on a skeleton of the major tracts. We find that SGL selects
FA metrics in the corticospinal tract and particualrly in the right
corticospinal tract as most important to ALS classification, confirming previous
findings\cite{van2011upper, toosy2003diffusion, sarica2014tractography,
sage2007quantitative, sage2009quantitative, karlsborg2004corticospinal,
ellis1999diffusion, cosottini2005diffusion, ciccarelli2009investigation,
abe2010voxel} and identifying exactly the portions of the brain that were
selected \emph{a priori} in the previous study from which we collected our
data\cite{sarica2017corticospinal}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.45\textwidth]{classification_probs.png}
    \includegraphics[width=0.45\textwidth]{classification_beta_bupu.png}

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.95\textwidth]{classification_tract_profiles.pdf}
    \caption{{\bf Model coefficients mirror FA differences}. The places along the
       length of the CST  where $\hat{\beta}$ coefficients for FA
       (dashed line, right axis) have large negative values correspond to the
       locations of substantial differences between the ALS (green) and control
       (orange) FA (shaded area indicates standard error of the mean).}
    \label{fig:class-profiles}
\end{figure}

\begin{figure}[!h]
    \centering
        \includegraphics[width=0.95\textwidth]{classification_subjects_profiles.pdf}
    \caption{{\bf Model mis-classifications correspond to features identified by
       the model}. The FA in CST of individuals that are mis-classified by the
       model is compared the group FA (with shaded are indicating standard error
       of the mean). False negative classifications (individuals that have ALS,
       but are classified as patients have) correspond to high FA either in the
       more inferior part of the CST}
    \label{fig:class-errors}
\end{figure}

>>>>

To test the performance of SGL in tractometry data in a continuous regression
task, we focus here on the prediction of biological age from the tractometry
data. Prediction of ``brain age'' is a commonly undertaken task. This is both
because it operates on a natural scale, with meaningful and easily understood
units, as well as because predictions of brain age, and deviations from accurate
prediction are diagnostic of overall brain health (for a recent review, see
\cite{Cole2019-rz}). The WH dataset used here contains data from 76 healthy
subjects, ranging between 6 years and 50 years of age
\cite{yeatman2014lifespan}. In this case, biological age was used as the
predicted variable ($y$ in equation \ref{eq:lm}). SGL was fit to
tractometry-extracted features: FA and MD in 20 major brain tracts, with each
tract divided into 100 nodes. To evaluate the fit of the model, we used a nested
cross-validation procedure. In this procedure, batches of subjects are held out.
For each batch (or fold), the model is fully fit without this data. Then, once
the parameters are fixed, the model is inverted to predict the ages of held out
subjects based on the linear coeffiecients and the static non-linearity.
This scheme automatically finds the right level of regularization and
fits the coefficients to the ill-posed linear model, while guarding against
overfitting. SGL accurately predicts the age of the subjects in this procedure,
with a mean absolute error of 3.6 years (Figure \ref{fig:regress-results}, left
panel). This is lower than the results of a recent study that predicted age in a
large sample, based on diffusion MRI features \cite{Richard2018-ux}.
Nevertheless, older subjects have higher residual variance, reflecting the
automatically-chosen log-transformation and implying that brain age becomes more
difficult to predict as we age chronologically (\ref{fig:regress-results}, right
panel). The model weights are distributed over many different tracts and dMRI
tissue properties (Figure \ref{fig:regress-beta} left). This demonstrates that
SGL is not coerced to produce overly sparse results when a more accurate model
requires a dense selection of features. Furthermore, looking closer at a
selection of tracts where high coefficients are found demonstrates that
diffusion properties (FA, in this case) are different in different age groups in
parts of the tracts where these higher coefficients are found (Figure
\ref{fig:regress-beta} right).

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{regression_residuals.png}
    \caption{{\bf Predicting age with tractometry and SGL.} Left: The predicted
    age of each individual (on the abscissa) and true age (on the ordinate),
    from the test splits (i.e., when each subject's data was held out in fitting
    the model); an accurate prediction falls close to the $y=x$ line (dashed).
    The mean absolute error in this case is 3.6 years and, the coefficient of
    determination $R^2=0.3$. Right: Standardized residuals (on the abscissa) as
    a function of the true age (on the ordinate). Predictions are generally more
    accurate for younger individuals.}
    \label{fig:regress-results}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.3\textwidth]{regression_beta_annotated.pdf}
    \includegraphics[width=0.67\textwidth]{regression_tract_profiles.pdf}
    \caption{{\bf Feature importance for predicting age from tractometry.} Left:
    A skeletonized display of the main brain tracts analyzed, with anterior
    facing up, and right hemisphere on the right. The $\hat{\beta}$ coefficients
    displayed in blue (negative) to red (positive) are for measurements of FA
    along the length of the tracts. The left cingulum cingulate (A) and forceps
    minor (B) are highlighted. Right: the FA (in shades of blue and green) and
    the $beta$ coefficients (dashed) in (A) left cingulum and (B) forceps
    minor.}
    \label{fig:regress-beta}
\end{figure}

@@

Using data from a previous study of the corticospinal tract (CST) profile and
ALS\cite{sarica2017corticospinal}, we tested the performance of SGL in a
classification setting. The previous study predicted ALS status with a mean
accuracy of 80\% using a random forest algorithm based on a priori selection of
features within the corticospinal tract. SGL delivers competitive predictive
performance (mean 93\% $\pm$ 2\% accuracy, 0.978 $\pm$ 0.006 ROC AUC) without
the need for a priori feature engineering. The results of the classification
prediction are shown in Fig~\ref{fig:class-results} with ``ground-truth'' ALS
status separated into columns, and predicted ALS status encoded by color. In
addition to this classification performance, SGL also identifies the white
matter tracts most important for ALS classification. The relative importance of
white matter features is captured in the $\beta$ coefficients from
Eq~\eqref{eq:sgl}. Fig~\ref{fig:class-profiles} depicts these coefficients along
the right CST, plotted over the FA values for the control and ALS subject
groups. We find that SGL selects
FA metrics in the corticospinal tract and particualrly in the right
corticospinal tract as most important to ALS classification, confirming previous
findings\cite{van2011upper, toosy2003diffusion, sarica2014tractography,
sage2007quantitative, sage2009quantitative, karlsborg2004corticospinal,
ellis1999diffusion, cosottini2005diffusion, ciccarelli2009investigation,
abe2010voxel} and identifying exactly the portions of the brain that were
selected \emph{a priori} in the previous study from which we collected our
data\cite{sarica2017corticospinal}.

We also find that mislabelled subjects correspond to outlying features in the
CST. Fig~\ref{fig:class-errors} depicts the group FA values along with
FA values of mislabelled subjects, two false negatives and one false-positive.
The false negative classifications have high FA in one of the two sections of
the CST where $\norm{\widehat{\beta}} > 0$ in Fig~\ref{fig:class-profiles}.
The false positive subject has an FA that oscillates between the two group
means. Thus, the SGL method fails comprehensibly.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.45\textwidth]{classification_probs.png}
    \includegraphics[width=0.45\textwidth]{classification_beta_bupu.png}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.95\textwidth]{classification_tract_profiles.pdf}
    \caption{{\bf Model coefficients mirror FA differences}. The places along the
       length of the CST  where $\hat{\beta}$ coefficients for FA
       (dashed line, right axis) have large negative values correspond to the
       locations of substantial differences between the ALS (green) and control
       (orange) FA (shaded area indicates standard error of the mean).}
    \label{fig:class-profiles}
\end{figure}

\begin{figure}[h]
    \centering
        \includegraphics[width=0.95\textwidth]{classification_subjects_profiles.pdf}
    \caption{{\bf Model mis-classifications correspond to features identified by
       the model}. The FA in CST of individuals that are mis-classified by the
       model is compared the group FA (with shaded are indicating standard error
       of the mean). False negative classifications (individuals that have ALS,
       but are classified as patients) correspond to high FA either in one of
       the two regions of large $\norm{\widehat{\beta}}$ in
       Fig~\ref{fig:class-profiles}. The false positive classification has
       an FA profile that oscillates between the two group means.
   }
    \label{fig:class-errors}
\end{figure}
